---
title: "How does littered trash differ in amount and composition, when being picked up from the greeneries next to a rural road in the same day, comparing three sections with different vicinities to the city?"

date: today
author:
  - name: Alexander Hüssy
    orcid: 0009-0007-7656-117X
    email: huessya@ethz.ch
    affiliation: 
      - name: ETH Zurich
        url: https://ethz.ch/de.html
format:
  html:
    toc: true

editor: visual
bibliography: references.bib
---

```{r}
#| label: load-packages

suppressMessages(library(tidyverse))
library(ggthemes)
library(ggridges)
library(gt)
```

# Introduction

Between villages and cities there are often rural roads, which are rarely frequented by pedestrians, but experience rather high amounts of car traffic. Usually these streets have a parallel running footway with green patches in between, and roughly one meter of grass on either side, bordering a field or forest. These patches can not be accessed by the usual road sweeping vehicle, so littered trash can only be removed from there by picking it up manually. Unlike the similar study by @rossi2023understanding, this case study focuses on this litter, an often forgotten area, where the solid waste collection system fails.

# Methods

To obtain the data the littered trash was picked manually using a simple grabber, and was afterwards counted and weighted with a kitchen scale. Parts of three rural streets were chosen, always with some distance to the closest settlement. They are all part of the same route, one would take to drive into Zürich city. @tbl-streets displays their data and noteworthy conditions. The streets were further divided into sections.

```{r}
#| label: tbl-streets
#| tbl-cap: An overview of the collected litter per km from the different streets

loc2 <- read_csv(here::here("data/final/location_overview.csv"),show_col_types = FALSE)

loc2 |> 
  group_by(street,length,notes) |> 
  summarise(mean_mass = mean(mass),
            mean_amount = mean(amount),
            .groups = 'drop') |> 
  gt() |> 
  fmt_number(columns = c(mean_mass, mean_amount),
             decimals = 0) |> 
   cols_label(
    length = "length [m]",
    mean_mass = "Mean Total Mass [g/km]",
    mean_amount = "Mean Total Amount [#/km]"
  )


```

# Results

@fig-sectionsA and @fig-sectionsB compare the mass and amount per km respectively in different sections of the same street, with the section "Zürichstrasse1" being closest to a village. This data agrees with the waste picker's suspicion, that closeness to a settlement correlates with the total amount of littered trash. However a clear outlier are cigarettes, which are roughly equally distributed everywhere (their pick-up-rate depends more on the accessibility, i.e. grass height).

```{r}
#| fig-cap: Comparison of different Sections of the same Street
#| label: fig-sectionsA

sec_z <- read_csv(here::here("data/final/section_comparison.csv"),show_col_types = FALSE)

ggplot(data = sec_z, mapping = aes(x = location, 
                                     y = mass, 
                                   fill = waste_category)) + 
  geom_bar(stat = "identity") + 
  scale_color_colorblind() +
  labs(title = "Mass comparison of different Sections", x = "Street Section", y = "Mass [g/km]", fill = "Waste Categories")

```

```{r}
#| fig-cap: Comparison of different Sections of the same Street
#| label: fig-sectionsB


ggplot(data = sec_z, mapping = aes(x = location, 
                                     y = amount, 
                                   fill = waste_category)) + 
  geom_bar(stat = "identity") + 
  scale_color_colorblind() +
  labs(title = "Amount comparison of different Sections", x = "Street Section", y = "Amount [#/km]", fill = "Waste Categories")

```

Comparing the different streets (@fig-streetsA, @fig-streetsB) there is no clear correlation with closeness to the city or state of the grass (Zürichstrasse is the furthest and Wittikonerstrasse(Pfaffhausen) is the closest to Zürich city).

```{r}
#| fig-cap: Comparison of different Streets
#| label: fig-streetsA

litter_mean <- read_csv(here::here("data/final/total_analysis.csv"),show_col_types = FALSE)

ggplot(data = litter_mean, mapping = aes(x = street, 
                                   y = mean_mass, 
                                   fill = waste_category)) + 
  geom_bar(stat = "identity") + 
  scale_color_colorblind() +
  labs(title = "Mass comparison of different Streets", x = "Street", y = "Mass [g/km]", fill = "Waste Categories")

```

```{r}
#| fig-cap: Comparison of different Streets
#| label: fig-streetsB

ggplot(data = litter_mean, mapping = aes(x = street, 
                                   y = mean_amount, 
                                   fill = waste_category)) + 
  geom_bar(stat = "identity") + 
  scale_color_colorblind() +
  labs(title = "Amount comparison of different Streets", x = "Street", y = "Amount [#/km]", fill = "Waste Categories")

```

# Conclusions

-   Similar to the findings of @schultz2013littering, the largest contributer (in numbers) is cigarettes

-   Closeness to settlements usually entails more littering, but not necessarily closeness to major cities

-   Trash in the green patches between streets need to be collected manually, perhaps with a system proposed by @powers1973positive
